How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

The contributions of this paper are two-fold:
\be
\item
The  XTREES planner, which has not been previously published. For demonstration purposes, we show XTREES
working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
can determine that some rows are preferred and some are to be avoided).  
\item An evaluation strategy used to comparatively assess planners. As  shown below, that strategy less us identify  inferior planners. With this strategy, other researchers could assess improvements or alternatives to the XTREE planner. 
\ei
The rest of this paper is structured as follows.
After describing our test data,
this paper  walks the reader through four methods that start
with a standard clustering algorithm, the  progresses on to XTREES. The second
half of the paper then presents an extensive evaluation of these methods using 
the Jureczko  and Siegmund data sets. This is followed by notes on related work
and validity.

@@

\begin{figure}
\includegraphics[width=2in]{_figs/twod.eps} 
\caption{Training examples (in gray), test examples (in red) and 
changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
date than the original test data
i.e. predictors trained on the training examples and
applied to the test examples should
work just as well, or better, on the changed test examples.}\label{fig:howxy}
\end{figure}
%XXX end

@@

 \begin{figure}[!t]
\scriptsize
\begin{tabular}{llllll}
  \hline
  \rowcolor{lightgray}
Project & Domain & Lang. & LOC & Features & Config\\\hline
Berkeley DB CE & Database & C & 219,811 & 18 & 2560\\
Berkeley DB JE & Database & Java & 42,596 & 32  & 400\\
Apache & Web Server & C & 230,277 & 9 & 192\\
SQLite & Database & C & 312,625 & 39 & 3,932,160\\
LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
\end{tabular}

@@

\begin{figure}[!t]
\includegraphics[width=1\linewidth]{_figs/BDBC.eps}
\caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
\end{figure}

@@

  \section{Test Data}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

@@

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei
  Note that all our methods have the two properties proposed in \tion{prelim}:
  \bi
  \item They are {\em local learners}; i.e. different test examples
  will be given plans that are specialized  to their particulars;
  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
  i.e. they change changing examples such that they move {\em closer} to the training data.
  \ei
The following descriptions of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

@@

@@

\subsubsection{Performance Measures}

For a planning system designed for business users, we seek two properties:
\bi
\item {\em Effectiveness:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  should  be  observed. 
\item {\em Succinctness:} given two plans with the same effectiveness, we prefer the shorter one.
Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
\ei
 In the following, we will show that XTREES and one other method (BIC) are equally effective
 but XTREES is more succinct (proposes changes to fewer features).

@@

{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}}\hline
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
1 &         XTREE &    0.49  &  0.13 & \quart{7}{25}{17}{115} \\\hline  
2 &      CD &    0.59  &  0.18 & \quart{15}{34}{36}{115} \\
2 &          BIC &    0.60  &  0.12 & \quart{24}{24}{38}{115} \\\hline  
3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115} \arrayrulecolor{black}\\\hline \end{tabular}}
\end{center}

@@

@@

\begin{figure}[t]
\begin{minipage}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{_figs/Deltas-Lucene.eps}
\label{fig:delta_ivy}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{_figs/Deltas-BDBJ.eps}
\label{fig:delta_apache}
\end{minipage}
\caption{Performance Comparison showing the number of times a certain feature requires to be changed. X-axis lists the various features that make up a data set. The y-axis contains the frequency of change in \%.}
\end{figure}

all the smoting and tuning running in train.

  The dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
\bi
\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data.
\item For the Jureczko data, we used the training and testing sets of \fig{j}. 
\ei
The 40 repeats defined above were repeated for each planner and the results processed as described in the last section.

\input{tex/deltas_avg.tex}
The results for each data, for different planners, are shown in \fig{jur} and \fig{conf}. All
\subsection{Discussion}

\section{Related work}
results suggest that, when clustering on independent variables of software project data,
	we are free to select from a wide range of 
	clustering methods.  Ganesan~\cite{div14} explored 
	different clustering methods for SE data using   effort and defect data from
	the PROMISE repository~\cite{promiserepo}.
	That study explored
	K-Means, mini-batch-K-Means, DBscan, EM, Ward, and the WHERE algorithm discussed
	below.
	Clusters were assessed via the performance of prediction 
	learned from the clusters by Random Forests (for defect prediction)
	and M5prime or linear regression (for effect estimation).  Ganesan found
	that the details of the clustering method were less important than ensuring that  a large number of clusters were generated.
	Accordingly, for clustering, we select a clustering method that generates a large
	number of small cluster
	($\sqrt{N}$ clusters
	from $N$ examples) that runs fasts, which has shown promise in prior work~\cite{Menzies2013}, and which ignores spurious dimensions (the last item is important since, in our experience, much SE data is ``noisey''; i.e. contains signals not associated with the target variable.
	ats to Validity}

>>>>

The contributions of this paper are two-fold:
\be
\item
The  XTREES planner, which has not been previously published. For demonstration purposes, we show XTREES
working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
can determine that some rows are preferred and some are to be avoided).  
\item An evaluation strategy used to comparatively assess planners. As  shown below, that strategy less us identify  inferior planners. With this strategy, other researchers could assess improvements or alternatives to the XTREE planner. 
\ee
In all, this paper explores and comparatively assesses four methods for learning planners (and XTREES is Method\#4).
As an historical note,  Method\#2 was previously proposed in one of our prior publications~\cite{me12c}. As shown below,
that method is a comparatively far less effective than XTREES. Hence, we   recant that prior work and, based on the results of this
paper, we now endorse  XTREES.

The rest of this paper is structured as follows.
After describing our test data,
this paper  walks the reader through four methods that start
with a standard clustering algorithm, the  progresses on to XTREES. The second
half of the paper then presents an extensive evaluation of these methods using 
the Jureczko  and Siegmund data sets. This is followed by notes on related work
and validity.

@@

\begin{figure}[!t]
  \includegraphics[width=2in,bb=0 0 100 100]{figs/aa.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Training examples (in gray), test examples (in red) and 
changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
date than the original test data
i.e. predictors trained on the training examples and
applied to the test examples should
work just as well, or better, on the changed test examples.}\label{fig:howxy}
\end{figure}

%XXX end

@@

 \begin{figure}[!t]
\scriptsize
\begin{tabular}{llllll}
  \hline
  \rowcolor{lightgray}
Project & Domain & Lang. & LOC & Features & Config\\\hline
BDBC: Berkeley DB   & Database & C & 219,811 & 18 & 2560\\
BDBJ: Berkeley DB   & Database & Java & 42,596 & 32  & 400\\
Apache & Web Server & C & 230,277 & 9 & 192\\
SQLite & Database & C & 312,625 & 39 & 3,932,160\\
LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
\end{tabular}

@@

\begin{figure}[!t]
\includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
\end{figure}

@@

  \section{Test Data}\label{sect:tesd}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

@@

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei
  Note that all our methods have the two properties proposed in \tion{prelim}:
  \bi
  \item They are {\em local learners}; i.e. different test examples
  will be given plans that are specialized  to their particulars;
  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
  i.e. they change changing examples such that they move {\em closer} to the training data.
  \ei

Our  description of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

@@

As an historical note,  Method\#2 was previously proposed in one of our prior publications~\cite{me12c}.

@@

@@

{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
1 &         XTREE &    0.49  &  0.13 & \quart{7}{25}{17}{115} \\\hline  
2 &      CD &    0.59  &  0.18 & \quart{15}{34}{36}{115} \\
2 &          BIC &    0.60  &  0.12 & \quart{24}{24}{38}{115} \\\hline  
3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
\end{center}

@@

\begin{figure}[t]
\begin{minipage}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{figs/Deltas-Lucene.eps}
\label{fig:delta_ivy}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{figs/Deltas-BDBJ.eps}
\label{fig:delta_apache}
\end{minipage}
\caption{Performance Comparison showing the number of times a certain feature requires to be changed. X-axis lists the various features that make up a data set. The y-axis contains the frequency of change in \%.}\label{fig:changed}
\end{figure}

@@

\subsubsection{Effectiveness Results}

For a planning system designed for business users, we seek two properties:
\bi
\item {\em Effectiveness:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  should  be  observed. 
\item {\em Succinctness:} given two plans with the same effectiveness, we prefer the shorter one.
Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
\ei

XTREES performs well in terms of  effectiveness. 
\fig{jur} and \fig{conf} show the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
\bi
\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
 all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
occurred in the {\em train} phase of \fig{work}.
\ei
The results show planning can be very effective-- some of the expected value of the reductions are very large:
\bi
\item
Down to 22\% of the original baseline in Ant of \fig{jur};
\item
Down to 22\% of the original baseline in X264  of \fig{conf};
\ei
In all these results,  with one exception, XTREES always is the top-ranked method and with a median improvement at least 10\% lower (i.e. better) than the next ranked method. The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
so XTREES can only reduce the expected values of the runtimes by 1\%.

From these results, measured in terms of effectiveness, we recommend supervised-based division of the data (as used in XTREE) over unsupervised clustering
(as used in the other methods). As to the very simple methods (CD and CD+FSS), these were proposed in prior publications
on this work~\cite{me12c}. We now   observe  that that prior work was poorly evaluated so the recommendations
of that prior paper  must now be rejected.

As to succinctness , that is discussed in the next section.

\subsubsection{Succinctness Results}

\input{tex/deltas_avg.tex}

\fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
The top plot of that figure reports results from one of the  Jureczko data sets ({\em lucene}) and the bottom
plot shows a Siegmund data set ({\em BDBJ}).

As predicted in the above discussion on Method 1 (CD), this method always proposes changes  to most features. Method3 (BIC) often proposes changes to many features, which 
would make its plan tedious to implement. 
Method 2 (CD+FS)
proposes less, due to its use of feature selection and, in the case of {\em BDBJ}, it fixates on changing just one feature (which
probably explains why CD+FS performs so poorly for {\em BDBJ} in \fig{conf}). Finally, Method4 (XTREES) rarely fixates
on changing one feature (only on {\em LUCENE}'s lines of code feature) and seems to prefer making many small changes,
all over the space of features.

\fig{changed} reports the changes just for   {\em lucene}) and {\em BDBJ}. \fig{types} shows the mean changes for features found by our
four methods (here, mean change is the average height of the histogram bars seen in, say, \fig{changed}). Overall, Method1 (CD) and
Method3 (BIC) make the most changes while XTREES and CD+FS make the fewest.

\section{Related work}
results suggest that, when clustering on independent variables of software project data,
	we are free to select from a wide range of 
	clustering methods.  Ganesan~\cite{div14} explored 
	different clustering methods for SE data using   effort and defect data from
	the PROMISE repository~\cite{promiserepo}.
	That study explored
	K-Means, mini-batch-K-Means, DBscan, EM, Ward, and the WHERE algorithm discussed
	below.
	Clusters were assessed via the performance of prediction 
	learned from the clusters by Random Forests (for defect prediction)
	and M5prime or linear regression (for effect estimation).  Ganesan found
	that the details of the clustering method were less important than ensuring that  a large number of clusters were generated.
	Accordingly, for clustering, we select a clustering method that generates a large
	number of small cluster
	($\sqrt{N}$ clusters
	from $N$ examples) that runs fasts, which has shown promise in prior work~\cite{Menzies2013}, and which ignores spurious dimensions (the last item is important since, in our experience, much SE data is ``noisey''; i.e. contains signals not associated with the target variable.
	ats to Validity
